---
title: "On the Origin of Citations"
author: "Rahul Subramanian"
date: "September 15, 2015"
output:
  word_document: default
  pdf_document:
    toc: yes
  html_document:
    fig_caption: yes
    fig_height: 8
    fig_width: 8
    highlight: espresso
    number_sections: yes
    self_contained: no
    theme: cerulean
    toc: yes
---

# Load the data

## Using read.delim
```{r load_data}
counts_raw <- read.delim("data/counts-raw.txt.gz")
counts_norm <- read.delim("data/counts-norm.txt.gz")
```

# Data exploration

What's the distribution of all articles of our data set?
```{r author_histogram, fig.cap="Figure 1: Number of Authors Per Article", echo=FALSE}
hist(counts_raw$authorsCount,main="Authors per paper", xlab = "# authors")
```

```{r, shares_histogram, fig.cap="Figure 2: Number of Facebook Shares Per Article", echo=FALSE}
hist(counts_raw$facebookShareCount, breaks = c(0,20,40,60,80,100,5000), main = "Facebook shares per article", xlab = "#of shares", xlim = c(0,100))
```

The average number of facebook shares per article is `r mean(counts_raw$facebookShareCount)` .

## dplyr

```{r}
library("dplyr")
```

```{r}
research <-filter(counts_raw, articleType == "Research Article")
```

```{r}
reserach_2006 <-filter(research, year == 2006)
nrow(research_2006)
```

```{r}
research_2006_fb <-filter(research, year == 2006, facebookCommentCount > 0)
nrow(research_2006_fb)
```

```{r}
research_2006_fb_tweet <-filter(research, year == 2006, facebookCommentCount > 0 | backtweetsCount > 0)
nrow(research_2006_fb_tweet)

```

```{r}
research_2006_fb_tweet_disease <-filter(research, year == 2006, facebookCommentCount > 0 | backtweetsCount > 0, grepl("Infectious Diseases", plosSubjectTags))
nrow(research_2006_fb_tweet_disease)
```

```{r}
article_info <-select(research, doi, pubDate, journal, title, articleType,authorsCount)
colnames(article_info)
```

```{r}
article_info <-select(research, doi:authorsCount)
colnames(article_info)
```

```{r}
metrics <- select(research, contains("Count"), -authorsCount, f1000Factor, wikipediaCites)
colnames(metrics)
```

```{r}
head(select(research, journal))
head(select(research, 3))

```

```{r}
slice(article_info,1:3)
```

```{r}
low_cite_big <- filter(research, year<2009, pdfDownloadsCount > 1000, mendeleyReadersCount>15,wosCountThru2011 < 10 )
low_cite <-select(low_cite_big, journal, title, year)
nrow(low_cite)
select(low_cite, title)
```

### Chaining commands with dplyr

pipe character is R: %>%

```{r}
facebook_2006 <-research %>%filter(year == 2006) %>%
  select(contains("facebook"))

head(facebook_2006)

facebook_2006 <-research %>%
  filter(year == 2006) %>%
  select(contains("facebook"))%>% 
  nrow

```


arrange works similar to  function order

```{r}
research %>% arrange(desc(authorsCount), desc(wosCountThru2011)) %>%
  select(authorsCount, wosCountThru2011) %>%
  slice(1:10)
```

```{r}
research %>% arrange(desc(wosCountThru2011), desc(title)) %>% 
  select(wosCountThru2011,title) %>%
  slice(1:3)
```


```{r}
research %>% arrange(desc(authorsCount)) %>%
  select(authorsCount,title,journal,plosSubjectTags) %>%
 slice(1:3)
```

### summarizing with dplyr
```{r}
research <-research %>% mutate(weeksSincePublished = daysSincePublished / 7, 
                               yearsSincePublished = weeksSincePublished / 52)

research %>% select(contains("Since")) %>% slice(1:10)
```


using summarize (Makes data frame)
```{r}
research %>% summarize(plos_mean = mean(plosCommentCount),
                       plos_sd = sd(plosCommentCount),
                       num = n())
```

### Using group_by
```{r}
research %>% group_by(journal,year) %>%
  summarize(tweets_mean = mean(backtweetsCount))
  
```

## Challenge
```{r}
tweets_per_journal <- research %>% group_by(journal) %>%
  summarise(num_art = n(),tweets_mean = mean(backtweetsCount), tweets_sem = sd(backtweetsCount)/sqrt(num_art))
tweets_per_journal
```

# ggplot2
```{r}
library("ggplot2")
```

```{r}
p<- ggplot(research, aes(x = pdfDownloadsCount,
                                           y = wosCountThru2011, color = journal))+
  geom_point(aes()) + geom_smooth()
p
```

Create a scatter plot with daysSincePublished mapped to the x-axis and wosCountThru2011 mapped to the y-axis. Include a loess fit of the data. Set the transparency level (alpha) of the points to 0.5 and color the points according to the journal where the article was published. Make the loess curve red.

```{r}
p<- ggplot(research, aes(x = daysSincePublished,
                                           y = wosCountThru2011, ))+
  geom_point(aes(color = journal), alpha = 0.5) + geom_smooth(color = "red")
p
```
Breaks at 1 and 3 and repalce with desired indices
### Using scales
```{r}
p<- ggplot(research, aes(x = log10(pdfDownloadsCount+1),
                                           y = log10(wosCountThru2011+1), ))+
  geom_point(aes(color = journal)) + 
  geom_smooth() + 
  scale_x_continuous(breaks = c(1,3), labels = c(10,1000)) +
  scale_y_continuous(breaks = c(1,3), labels = c(10,1000), 
                     limits = c(1,3))
p 
```

different color options
```{r}
p + scale_color_grey()
p + scale_color_manual(values = c("red","green", "blue", "orange", "pink", "yellow", "purple"))
```
built in packages

```{r}
library("RColorBrewer")
display.brewer.all(type = "qual")
```

```{r}

p + scale_color_brewer(palette  = "Dark2",
                       labels = 1:7, name = "PLOS")

```

Update the plot to use a square root transformation instead of log10. Also color the points using the ColorBrewer palette “Accent”.
```{r}
p<- ggplot(research, aes(x = sqrt(pdfDownloadsCount),
                                           y = sqrt(wosCountThru2011), ))+
  geom_point(aes(color = journal)) + 
  geom_smooth() + 
 
                      scale_color_brewer(palette  = "Accent")
p 
```

```{r}
p<- ggplot(research, aes(x = sqrt(pdfDownloadsCount+1),
                                           y = sqrt(wosCountThru2011+1), ))+
  geom_point(aes(color = journal)) + 
  geom_smooth() + scale_color_brewer(palette  = "Accent")

```

### Using facets to make subplots
```{r}
p<- ggplot(research, aes(x = sqrt(pdfDownloadsCount),
                                           y = sqrt(wosCountThru2011), ))+
  geom_point(aes(color = journal)) + 
  geom_smooth() + 
 
                      scale_color_brewer(palette  = "Accent")
p + facet_wrap(~journal, ncol = 2) 
```

using facet grid

```{r}
research <- mutate(research, immuno = grepl("Immunology", plosSubjectTags)) 
p<- ggplot(research, aes(x = sqrt(pdfDownloadsCount),
                                           y = sqrt(wosCountThru2011), ))+
  geom_point(aes(color = journal)) + 
  geom_smooth() + 
 
                      scale_color_brewer(palette  = "Accent")
p + 
facet_grid(journal~immuno)

```


### Using different geoms
```{r}
p<- ggplot(research, aes(x = journal,
                                           y = sqrt(wosCountThru2011), ))+
  geom_boxplot()
p
```
making a barplot
from earlier:
```{r}
tweets_per_journal <- research %>% group_by(journal) %>%
  summarise(num_art = n(),tweets_mean = mean(backtweetsCount), tweets_sem = sd(backtweetsCount)/sqrt(num_art))
tweets_per_journal
```
now:
```{r}
tweets_bar <- ggplot(tweets_per_journal, aes(x= journal, y = tweets_mean)) +
  geom_bar(stat = "identity") + 
  geom_errorbar(aes(ymin = tweets_mean - tweets_sem, ymax = tweets_mean + tweets_sem))+ geom_text(aes(label = num_art), hjust = 0, vjust = 0)
  geom_bar(stat = "identity")

tweets_bar
```

Modify the dplyr code above to calculate the mean, SEM, and sample size of the number of article tweets per journal and per year. Use facet_wrap to make a separate subplot per year.

```{r}
tweets_per_journal <- research %>% group_by(journal, year) %>%
  summarise(num_art = n(),tweets_mean = mean(backtweetsCount), tweets_sem = sd(backtweetsCount)/sqrt(num_art))
tweets_per_journal
```

```{r}
tweets_bar <- ggplot(tweets_per_journal, aes(x= journal, y = tweets_mean)) +
  geom_bar(stat = "identity") + 
  geom_errorbar(aes(ymin = tweets_mean - tweets_sem, ymax = tweets_mean + tweets_sem)) + facet_wrap(~year)
```

### Customizing the plot
```{r}
tweets_bar + labs(title = "Mean tweests per journal per year", x = "Journal", y = "Mean number of tweets")+ theme_minimal()
```

```{r}
tweets_bar + theme_bw()
tweets_bar + theme_classic()

```

